---
title: "Spatial Analysis - Violent Crimes in Boston"
author: "Jonathan Schierbaum"
date: "SPRING 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <- c("mgcv", "dplyr", "sf", 
              "ggplot2", "ggmap", "gifski", 
              "mgcViz", "viridis", "colorspace", 
              "RColorBrewer", "grid", "gridGraphics", 
              "gridExtra")

missing <- packages[!(packages %in% installed.packages())]

if(length(missing) > 0) install.packages(missing) 

invisible(suppressMessages(
  lapply(packages, library, character.only = TRUE)))
```

## Load pre-constructed data objects For faster analysis

```{r load_data_objs}
fp_raw_data <- "~/GitHub/project-spatial-poisson-crime/raw_data/"

fp_data_obj <- "~/GitHub/project-spatial-poisson-crime/data_objs/"

fp_figures <- "~/GitHub/project-spatial-poisson-crime/figures/"

boundary.sf <- read_sf(dsn = paste0(fp_raw_data, "City_of_Boston_Boundary"), 
                       layer = "City_of_Boston_Boundary")

boundary.sf.4326 <- st_transform(boundary.sf, crs = 4326)

project_crs <- st_crs(boundary.sf.4326)


load(paste0(fp_data_obj, "boston_regression_objects.RData")) # Regression model objects

load(paste0(fp_data_obj, "boston_predictions.RData"))        # data grids WITH predictions

load(paste0(fp_data_obj, "boston_predictions_sf.RData"))     # data grids WITH predictions AND spatial embed

load(paste0(fp_data_obj, "boston_spatial_project_r1.RData")) # mapping objects and raw data
```

## EDA plots (observed counts)
```{r observed_plots}
X_max  <- max(st_coordinates(bounded_map.15to21)[,1])
X_min  <- min(st_coordinates(bounded_map.15to21)[,1])
Y_max  <- max(st_coordinates(bounded_map.15to21)[,2])
Y_min  <- min(st_coordinates(bounded_map.15to21)[,2])

basemap.boston <- get_stamenmap(bbox = c(X_min,Y_min,X_max,Y_max),
                    maptype = "terrain-lines", zoom = 13)

scale_lim.low  <- min(project.grid.all$n)
scale_lim.high <- max(project.grid.all$n)
scale_lim.high <- 480  # produces a slightly nicer plot.

### GIF OF OBSERVED
width_in <-  7
height_in <- 5
reso <- 225

png_fp <- file.path(tempdir(), "frame%03d.png")
png(png_fp, width = (width_in * reso), height = (height_in * reso), res = reso)
par(ask = FALSE)

ggmap(basemap.boston) + 
        geom_sf(aes(fill = n), data = project.grid.all, alpha = 0.75, inherit.aes = F) + 
        scale_fill_viridis(breaks = seq(scale_lim.low, scale_lim.high, length.out = 13), 
                           limits = c(scale_lim.low, scale_lim.high), 
                           option = "E", name = "Annual violent \ncrimes") + 
        ggtitle("Violent Crimes in Boston: 2015-2021") + 
        guides(fill = guide_legend(override.aes = list(size = 1))) + 
        guides(fill = guide_legend(reverse = T)) + 
        theme(axis.title = element_blank())   

scale_lim.low  <- min(c(project.grid.15$n, project.grid.16$n, project.grid.17$n, 
                        project.grid.18$n, project.grid.19$n, project.grid.20$n, 
                        project.grid.21$n))

scale_lim.high <- max(c(project.grid.15$n, project.grid.16$n, project.grid.17$n, 
                        project.grid.18$n, project.grid.19$n, project.grid.20$n, 
                        project.grid.21$n))

      
ggmap(basemap.boston) + 
        geom_sf(aes(fill = n), data = project.grid.15, alpha = 0.75, inherit.aes = F) + 
        scale_fill_viridis(breaks = seq(scale_lim.low, scale_lim.high, length.out = 13), 
                           limits = c(scale_lim.low, scale_lim.high), 
                           option = "E", name = "Annual violent \ncrimes") + 
        ggtitle("Violent Crimes in Boston: 2015") + 
        guides(fill = guide_legend(override.aes = list(size = 1))) + 
        guides(fill = guide_legend(reverse = T)) + 
        theme(axis.title = element_blank())   

ggmap(basemap.boston) +       
        geom_sf(aes(fill = n), data = project.grid.16, alpha = 0.75, inherit.aes = F) + 
        scale_fill_viridis(breaks = seq(scale_lim.low, scale_lim.high, length.out = 13), 
                           limits = c(scale_lim.low, scale_lim.high), 
                           option = "E", name = "Annual violent \ncrimes") + 
        ggtitle("Violent Crimes in Boston: 2016") + 
        guides(fill = guide_legend(override.aes = list(size = 1))) + 
        guides(fill = guide_legend(reverse = T)) + 
        theme(axis.title = element_blank())   

ggmap(basemap.boston) +       
        geom_sf(aes(fill = n), data = project.grid.17, alpha = 0.75, inherit.aes = F) + 
        scale_fill_viridis(breaks = seq(scale_lim.low, scale_lim.high, length.out = 13), 
                           limits = c(scale_lim.low, scale_lim.high), 
                           option = "E", name = "Annual violent \ncrimes") + 
        ggtitle("Violent Crimes in Boston: 2017") + 
        guides(fill = guide_legend(override.aes = list(size = 1))) + 
        guides(fill = guide_legend(reverse = T)) + 
        theme(axis.title = element_blank())   

ggmap(basemap.boston) +       
        geom_sf(aes(fill = n), data = project.grid.18, alpha = 0.75, inherit.aes = F) + 
        scale_fill_viridis(breaks = seq(scale_lim.low, scale_lim.high, length.out = 13), 
                           limits = c(scale_lim.low, scale_lim.high), 
                           option = "E", name = "Annual violent \ncrimes") + 
        ggtitle("Violent Crimes in Boston: 2018") + 
        guides(fill = guide_legend(override.aes = list(size = 1))) + 
        guides(fill = guide_legend(reverse = T)) + 
        theme(axis.title = element_blank())   

ggmap(basemap.boston) +       
        geom_sf(aes(fill = n), data = project.grid.19, alpha = 0.75, inherit.aes = F) + 
        scale_fill_viridis(breaks = seq(scale_lim.low, scale_lim.high, length.out = 13), 
                           limits = c(scale_lim.low, scale_lim.high), 
                           option = "E", name = "Annual violent \ncrimes") + 
        ggtitle("Violent Crimes in Boston: 2019") + 
        guides(fill = guide_legend(override.aes = list(size = 1))) + 
        guides(fill = guide_legend(reverse = T)) + 
       theme(axis.title = element_blank())   
      
ggmap(basemap.boston) +       
        geom_sf(aes(fill = n), data = project.grid.20, alpha = 0.75, inherit.aes = F) + 
        scale_fill_viridis(breaks = seq(scale_lim.low, scale_lim.high, length.out = 13), 
                           limits = c(scale_lim.low, scale_lim.high), 
                           option = "E", name = "Annual violent \ncrimes") + 
        ggtitle("Violent Crimes in Boston: 2020") + 
        guides(fill = guide_legend(override.aes = list(size = 1))) + 
        guides(fill = guide_legend(reverse = T)) + 
        theme(axis.title = element_blank())   

ggmap(basemap.boston) +       
        geom_sf(aes(fill = n), data = project.grid.21, alpha = 0.75, inherit.aes = F) + 
        scale_fill_viridis(breaks = seq(scale_lim.low, scale_lim.high, length.out = 13), 
                           limits = c(scale_lim.low, scale_lim.high), 
                           option = "E", name = "Annual violent \ncrimes") + 
        ggtitle("Violent Crimes in Boston: 2021") + 
        guides(fill = guide_legend(override.aes = list(size = 1))) + 
        guides(fill = guide_legend(reverse = T)) + 
        theme(axis.title = element_blank())   

scale_lim.low  <- min(project.grid.all$n)
scale_lim.high <- max(project.grid.all$n)
scale_lim.high <- 480  # produces a slightly nicer plot.

ggmap(basemap.boston) +  
        geom_sf(aes(fill = n), data = project.grid.all, alpha = 0.75, inherit.aes = F) + 
        scale_fill_viridis(breaks = seq(scale_lim.low, scale_lim.high, length.out = 13), 
                           limits = c(scale_lim.low, scale_lim.high), 
                           option = "E", name = "Annual violent \ncrimes") + 
        ggtitle("Violent Crimes in Boston: 2015-2021") + 
        guides(fill = guide_legend(override.aes = list(size = 1))) + 
        guides(fill = guide_legend(reverse = T)) + 
        theme(axis.title = element_blank())   

dev.off()

png_files <- sprintf(png_fp, 1:9)
gif_fp <- paste0(fp_figures, "observed_violent_crimes.gif")
gifski(png_files, gif_fp, delay = 2)
unlink(png_files)
utils::browseURL(gif_fp)
```


## Model selection and diagnostics

```{r model_selection_diagnostics}
### MODEL MATRIX FOR ALL VIOLENT CRIME ###
modelall.df <- as.data.frame(st_drop_geometry(project.grid.all[, 2:4]))

names(modelall.df) <- c("x", "y", "crimes") 

for (i in 1:nrow(modelall.df)){
  if (is.na(modelall.df[i,"crimes"])) {modelall.df[i,"crimes"] <- 0}  
  else {modelall.df[i,"crimes"] <- modelall.df[i,"crimes"]}  
}

### BUILD DIFFERENT MODELS ###

# k: set an upper limit on the ﬂexibility of a smooth term
model1 <- gam(crimes ~ s(x, y, k = 300), family = nb, method = "REML", data = modelall.df)

model2 <- gam(crimes ~ s(x, y, k = 300), family = poisson(link = "log"), method = "REML", data = modelall.df)

model3 <- gam(crimes ~ s(x, y, k = 300), family = quasipoisson(link = "log"), method = "REML", data = modelall.df)

model4 <- gam(crimes ~ s(x, y, k = 300), family = ziP(), data = modelall.df)

model5 <- gam(crimes ~ s(x, k = 25, bs = "cr")+ s(y, k = 25, bs = "cr") + ti(x, y, k = 25), family = nb, method = "REML", data = modelall.df)

model6 <- gam(crimes ~ te(x, y, k = 25), family = nb, method = "REML", data = modelall.df)

### DIAGNOSE DIFFERENT MODELS ###
check.gamViz(getViz(model1))
check.gamViz(getViz(model2))
check.gamViz(getViz(model3))
check.gamViz(getViz(model4))
check.gamViz(getViz(model5))
check.gamViz(getViz(model6))

### OTHER DIAGNOSES ###
AIC(model1, model2, model3, model4, model5, model6)

### CHECK RATE OF VARIANCE INCREASE ### 
e <- residuals(model1)
fitted1 <- fitted(model1)
lm(log(e^2) ~ log(fitted1))
# looks ok....if log(fitted) = 2, then the variance of the data increases with the square of the mean.

### CHECK RESIDUALS ###
fitted2  <- fitted(model2)
fitted3  <- fitted(model3)
fitted5  <- fitted(model5)
observed.all <- project.grid.all$n

sum_resids_1 <- sum((fitted1 - observed.all)^2) # RSS
sum_resids_2 <- sum((fitted2 - observed.all)^2)
sum_resids_3 <- sum((fitted3 - observed.all)^2)
sum_resids_5 <- sum((fitted5 - observed.all)^2)
```

```{r diagnostic_plot}
### COMPARE RESPONSE TYPES
model1.viz <- getViz(model1)
model2.viz <- getViz(model2)

check.gamViz(model1.viz)
check.gamViz(model2.viz)


# QQ plots:  use direct simulation. For each replicate, data are simulated from the fitted model, and the corresponding residuals computed.
viz1 <- qq.gamViz(model1.viz, rep = 25, showReps = T, CI = "none", a.qqpoi = list("shape" = 19), a.replin = list("alpha" = 0.2)) + labs(title = "Response: Negative Binomial Family") +
xlab("theoretical quantiles")

  viz2 <- qq.gamViz(model2.viz, rep = 25, showReps = T, CI = "none", a.qqpoi = list("shape" = 19), a.replin = list("alpha" = 0.2)) + labs(title = "Response: Poisson Family") +
xlab("theoretical quantiles")

width_in <-  10
height_in <- 5
reso <- 225

png(file=paste0(fp_figures,"family_diagnostics_qq", ".png"), width = width_in, height = height_in, unit = "in", res = reso, antialias = "cleartype")

gridPrint(viz1,viz2, ncol = 2)

dev.off()

# histogram of residuals
resids.df1 <- as.data.frame(cbind(1:length(residuals.gamViz(model1.viz)), residuals.gamViz(model1.viz)))
resids.df2 <- as.data.frame(cbind(1:length(residuals.gamViz(model2.viz)), residuals.gamViz(model2.viz)))

names(resids.df1) <- names(resids.df2) <- c("id", "deviance_resids")

viz3 <- ggplot(data = resids.df1, aes(deviance_resids)) +                         
          geom_histogram(alpha = 0.5) + 
          geom_vline(aes(xintercept=median(deviance_resids)),                 
                     color = "red", linetype = "dashed", size = 1 ) + 
          xlab("deviance residuals") + 
          ylab("frequency") +
          labs(title = "Response: Negative Binomial Family")

viz4 <- ggplot(data = resids.df2, aes(deviance_resids)) +                         
          geom_histogram(alpha = 0.5) + 
          geom_vline(aes(xintercept=median(deviance_resids)),                 
                     color = "red", linetype = "dashed", size = 1) + 
          xlab("deviance residuals") + 
          ylab("frequency") + 
          labs(title = "Response: Poisson Family")

width_in <-  10
height_in <- 5
reso <- 225

png(file=paste0(fp_figures,"family_diagnostics_resids_hist", ".png"), width = width_in, height = height_in, unit = "in", res = reso, antialias = "cleartype")

grid.arrange(viz3, viz4, ncol = 2)

dev.off()

# smoothing term

plot(sm(model1.viz, 1)) + labs(subtitle = "Smooth effect, DOF = 155.53") + labs(title = "Response: Negative Binomial Family") + xlab("longitude") + ylab("latitude")

viz5 <- grid.grab()

dev.off()

plot(sm(model2.viz, 1)) + labs(subtitle = "Smooth effect, DOF = 273.55") + labs(title="Response: Poisson Family") + xlab("longitude") + ylab("latitude")

viz6 <- grid.grab()

dev.off()

width_in <-  10
height_in <- 5.5
reso <- 225

png(file=paste0(fp_figures,"family_diagnostics_smoothing", ".png"), width = width_in, height = height_in, unit = "in", res = reso, antialias = "cleartype")

grid.arrange(viz5, viz6, ncol = 2)

dev.off()

### COMPARE SPLINE TYPES

model5.viz <- getViz(model5)

check.gamViz(model1.viz)
check.gamViz(model5.viz)

plot.gamViz(model1.viz, residuals = T, allTerms = T)
plot.gamViz(model5.viz, residuals = T, allTerms = T)
```

### Final model selection: train/test set prediction performance

```{r train_test_competition, eval= FALSE}
### PREDICTION CHECK OF TOP THREE MODELS ###
### HOLD OUT 20% OF DATA, PREDICT, REPEAT.

n_iter <- 100

pred_results <- matrix(0, nrow = n_iter, ncol = 3)

for (i in 1:n_iter){ 
  idx <- sample(1:nrow(modelall.df), size = ceiling(nrow(modelall.df) * .20))

  train.df <- modelall.df[-idx,]
  test.df  <- modelall.df[idx,]

  model.1 <- gam(crimes ~ s(x, y, k = 300), family = nb, method = "REML", data = train.df)

  model.3 <- gam(crimes ~ s(x, y, k = 300), family = quasipoisson(link = "log"), method = "REML", data = train.df)

  model.5 <- gam(crimes ~ s(x, k = 25, bs = "cr")+ s(y, k = 25, bs = "cr") + ti(x, y, k = 25), family = nb, method = "REML", data = train.df)

  response.preds1 <- as.vector(predict(model.1, newdata = test.df[,1:2], type = "response"))
  response.preds3 <- as.vector(predict(model.3, newdata = test.df[,1:2], type = "response"))
  response.preds5 <- as.vector(predict(model.5, newdata = test.df[,1:2], type = "response"))

  observed1 <- test.df$crimes

  check1 <- sum((response.preds1 - observed1)^2)
  check3 <- sum((response.preds3 - observed1)^2)
  check5 <- sum((response.preds5 - observed1)^2)
  
  pred_results[i, 1] <- check1
  pred_results[i, 2] <- check3
  pred_results[i, 3] <- check5
}
                       #   model1   model3   model5
colMeans(pred_results) #  148659.2 147667.0 146370.7


rm(model.1, model.2, model.3, test.df, train.df)
```

## Spatial patterning in residuals (plots)

```{r resids_spatial_dependence}
# note: A Gaussian GAM with identity link is just an ordinary linear model. In this situation, response residuals, deviance residuals, and pearson residuals are the same.


years <- c("15", "16", "17", "18", "19", "20", "21")


residuals.array <- array(matrix(0, nrow = nrow(modelall.df), ncol = (ncol(modelall.df) + 5)), dim = c(nrow(modelall.df), (ncol(modelall.df) + 5), length(years)))

for (i in years){ 

  resids.response       <- residuals(eval(as.name(paste0("model", i))), type = "response")        # Response residuals: raw residuals (data minus fitted values)



  resids.deviance       <- residuals(eval(as.name(paste0("model", i))), type = "deviance")        # Deviance residuals: residuals defined by model family



  resids.pearson        <- residuals(eval(as.name(paste0("model", i))), type = "pearson")         # Pearson residuals: raw residuals divided by the standard deviation of the data 



  resids.pearson.scaled <- residuals(eval(as.name(paste0("model", i))), type = "scaled.pearson")  # Scaled Pearson residuals: raw residuals divided by the standard deviation of the data 
                                                                                                  # according to the model mean variance relationship and estimated scale parameter.


  resids.working        <- residuals(eval(as.name(paste0("model", i))), type = "working")         # Working residuals: residuals returned from model fitting at convergence


  resids.all            <- cbind(eval(as.name(paste0("model", i, ".df"))), 
                                 resids.response, resids.deviance, resids.pearson, 
                                 resids.pearson.scaled, resids.working)  

  counter               <- as.numeric(which(years == i))
  
  residuals.array[ , , counter] <- as.matrix(resids.all)
  }

resids.15 <- as.data.frame(residuals.array[ , , which(years == "15")])

names(resids.15) <- c("lon", "lat", "crimes", "response.resids", "deviance.resids", 
                      "pearson.resids", "pearson.scaled.resids", "working.resids")

resids.15 <- st_as_sf(resids.15, coords = c("lon", "lat"), crs = st_crs(4326))



resids.16 <- as.data.frame(residuals.array[ , , which(years == "16")])

names(resids.16) <- c("lon", "lat", "crimes", "response.resids", "deviance.resids", 
                      "pearson.resids", "pearson.scaled.resids", "working.resids")

resids.16 <- st_as_sf(resids.16, coords = c("lon", "lat"), crs = st_crs(4326))



resids.17 <- as.data.frame(residuals.array[ , , which(years == "17")])

names(resids.17) <- c("lon", "lat", "crimes", "response.resids", "deviance.resids", 
                      "pearson.resids", "pearson.scaled.resids", "working.resids")

resids.17 <- st_as_sf(resids.17, coords = c("lon", "lat"), crs = st_crs(4326))



resids.18 <- as.data.frame(residuals.array[ , , which(years == "18")])

names(resids.18) <- c("lon", "lat", "crimes", "response.resids", "deviance.resids", 
                      "pearson.resids", "pearson.scaled.resids", "working.resids")

resids.18 <- st_as_sf(resids.18, coords = c("lon", "lat"), crs = st_crs(4326))



resids.19 <- as.data.frame(residuals.array[ , , which(years == "19")])

names(resids.19) <- c("lon", "lat", "crimes", "response.resids", "deviance.resids", 
                      "pearson.resids", "pearson.scaled.resids", "working.resids")

resids.19 <- st_as_sf(resids.19, coords = c("lon", "lat"), crs = st_crs(4326))



resids.20 <- as.data.frame(residuals.array[ , , which(years == "20")])

names(resids.20) <- c("lon", "lat", "crimes", "response.resids", "deviance.resids", 
                      "pearson.resids", "pearson.scaled.resids", "working.resids")

resids.20 <- st_as_sf(resids.20, coords = c("lon", "lat"), crs = st_crs(4326))


resids.21 <- as.data.frame(residuals.array[ , , which(years == "21")])

names(resids.21) <- c("lon", "lat", "crimes", "response.resids", "deviance.resids", 
                      "pearson.resids", "pearson.scaled.resids", "working.resids")

resids.21 <- st_as_sf(resids.21, coords = c("lon", "lat"), crs = st_crs(4326))



width_in <-  7
height_in <- 5
reso <- 225

png_fp <- file.path(tempdir(), "frame%03d.png")
png(png_fp, width = (width_in * reso), height = (height_in * reso), res = reso)
par(ask = FALSE)

ggmap(basemap.boston) + 
        geom_sf(aes(color = pearson.resids), data = resids.15, alpha = 0.75, inherit.aes = F) + 
        scale_color_gradient2("Pearson \nresiduals", low = "blue", mid = "darkgray", high = "red") + 
  ggtitle("Spatial Patterning Within (Pearson) Residuals: 2015") + 
  theme(axis.title = element_blank())   

ggmap(basemap.boston) + 
        geom_sf(aes(color = pearson.resids), data = resids.16, alpha = 0.75, inherit.aes = F) + 
        scale_color_gradient2("Pearson \nresiduals", low = "blue", mid = "darkgray", high = "red") + 
  ggtitle("Spatial Patterning Within (Pearson) Residuals: 2016") + 
  theme(axis.title = element_blank())

ggmap(basemap.boston) + 
        geom_sf(aes(color = pearson.resids), data = resids.17, alpha = 0.75, inherit.aes = F) + 
        scale_color_gradient2("Pearson \nresiduals", low = "blue", mid = "darkgray", high = "red") + 
  ggtitle("Spatial Patterning Within (Pearson) Residuals: 2017") + 
  theme(axis.title = element_blank())

ggmap(basemap.boston) + 
        geom_sf(aes(color = pearson.resids), data = resids.18, alpha = 0.75, inherit.aes = F) + 
        scale_color_gradient2("Pearson \nresiduals", low = "blue", mid = "darkgray", high = "red") + 
  ggtitle("Spatial Patterning Within (Pearson) Residuals: 2018") + 
  theme(axis.title = element_blank())

ggmap(basemap.boston) + 
        geom_sf(aes(color = pearson.resids), data = resids.19, alpha = 0.75, inherit.aes = F) + 
        scale_color_gradient2("Pearson \nresiduals", low = "blue", mid = "darkgray", high = "red") + 
  ggtitle("Spatial Patterning Within (Pearson) Residuals: 2019") + 
  theme(axis.title = element_blank())

ggmap(basemap.boston) + 
        geom_sf(aes(color = pearson.resids), data = resids.20, alpha = 0.75, inherit.aes = F) + 
        scale_color_gradient2("Pearson \nresiduals", low = "blue", mid = "darkgray", high = "red") + 
  ggtitle("Spatial Patterning Within (Pearson) Residuals: 2020") + 
  theme(axis.title = element_blank())

ggmap(basemap.boston) + 
        geom_sf(aes(color = pearson.resids), data = resids.21, alpha = 0.75, inherit.aes = F) + 
        scale_color_gradient2("Pearson \nresiduals", low = "blue", mid = "darkgray", high = "red") + 
  ggtitle("Spatial Patterning Within (Pearson) Residuals: 2021") + 
  theme(axis.title = element_blank())

dev.off()

png_files <- sprintf(png_fp, 1:7)
gif_fp <- paste0(fp_figures, "residual_spatial_patterning.gif")
gifski(png_files, gif_fp, delay = 2)
unlink(png_files)
utils::browseURL(gif_fp)
```

## Interpolate crime count across grid of points

```{r prediction_data_objs}

### READY REFERENCE DATASETS BY YEAR ###
model15.df <- as.data.frame(st_drop_geometry(project.grid.15[, 2:4]))

names(model15.df) <- c("x", "y", "crimes") 

for (i in 1:nrow(model15.df)){
  if (is.na(model15.df[i, "crimes"])) {model15.df[i, "crimes"] <- 0}  
  else {model15.df[i, "crimes"] <- model15.df[i, "crimes"]}  
  }

model16.df <- as.data.frame(st_drop_geometry(project.grid.16[,2:4]))

names(model16.df) <- c("x", "y", "crimes") 

for (i in 1:nrow(model16.df)){
  if (is.na(model16.df[i, "crimes"])) {model16.df[i, "crimes"] <- 0}  
  else {model16.df[i, "crimes"] <- model16.df[i, "crimes"]}  
  }

model17.df <- as.data.frame(st_drop_geometry(project.grid.17[, 2:4]))

names(model17.df) <- c("x", "y", "crimes") 

for (i in 1:nrow(model17.df)){
  if (is.na(model17.df[i, "crimes"])) {model17.df[i, "crimes"] <- 0}  
  else {model17.df[i, "crimes"] <- model17.df[i, "crimes"]}  
  }

model18.df <- as.data.frame(st_drop_geometry(project.grid.18[, 2:4]))

names(model18.df) <- c("x", "y", "crimes") 

for (i in 1:nrow(model18.df)){
  if (is.na(model18.df[i, "crimes"])) {model18.df[i, "crimes"] <- 0}  
  else {model18.df[i, "crimes"] <- model18.df[i, "crimes"]}  
  }

model19.df <- as.data.frame(st_drop_geometry(project.grid.19[, 2:4]))

names(model19.df) <- c("x", "y", "crimes") 

for (i in 1:nrow(model19.df)){
  if (is.na(model19.df[i, "crimes"])) {model19.df[i, "crimes"] <- 0}  
  else {model19.df[i, "crimes"] <- model19.df[i, "crimes"]}  
  }

model20.df <- as.data.frame(st_drop_geometry(project.grid.20[, 2:4]))

names(model20.df) <- c("x", "y", "crimes") 

for (i in 1:nrow(model20.df)){
  if (is.na(model20.df[i, "crimes"])) {model20.df[i, "crimes"] <- 0}  
  else {model20.df[i, "crimes"] <- model20.df[i, "crimes"]}  
  }

model21.df <- as.data.frame(st_drop_geometry(project.grid.21[, 2:4]))

names(model21.df) <- c("x", "y", "crimes") 

for (i in 1:nrow(model21.df)){
  if (is.na(model21.df[i, "crimes"])) {model21.df[i, "crimes"] <- 0}  
  else {model21.df[i, "crimes"] <- model21.df[i, "crimes"]}  
}

modelall.df <- as.data.frame(st_drop_geometry(project.grid.all[, 2:4]))

names(modelall.df) <- c("x", "y", "crimes") 

for (i in 1:nrow(modelall.df)){
  if (is.na(modelall.df[i, "crimes"])) {modelall.df[i, "crimes"] <- 0}  
  else {modelall.df[i, "crimes"] <- modelall.df[i, "crimes"]}  
}

### USE BEST MODEL AND RUN BY YEAR
model15   <- gam(crimes ~ s(x, k = 25, bs = "cr") + s(y, k = 25, bs = "cr") + ti(y, x, k = 25), family = nb, method="REML", data = model15.df)
model16   <- gam(crimes ~ s(x, k = 25, bs = "cr") + s(y, k = 25, bs = "cr") + ti(y, x, k = 25), family = nb, method="REML", data = model16.df)
model17   <- gam(crimes ~ s(x, k = 25, bs = "cr") + s(y, k = 25, bs = "cr") + ti(y, x, k = 25), family = nb, method="REML", data = model17.df)
model18   <- gam(crimes ~ s(x, k = 25, bs = "cr") + s(y, k = 25, bs = "cr") + ti(y, x, k = 25), family = nb, method="REML", data = model18.df)
model19   <- gam(crimes ~ s(x, k = 25, bs = "cr") + s(y, k = 25, bs = "cr") + ti(y, x, k = 25), family = nb, method="REML", data = model19.df)
model20   <- gam(crimes ~ s(x, k = 25, bs = "cr") + s(y, k = 25, bs = "cr") + ti(y, x, k = 25), family = nb, method="REML", data = model20.df)
model21   <- gam(crimes ~ s(x, k = 25, bs = "cr") + s(y, k = 25, bs = "cr") + ti(y, x, k = 25), family = nb, method="REML", data = model21.df)
model.all <- gam(crimes ~ s(x, k = 25, bs = "cr") + s(y, k = 25, bs = "cr") + ti(y, x, k = 25), family = nb, method="REML", data = modelall.df)


### MAKE GRID TO PREDICT ACROSS ###
# X_max  <- max(c(st_bbox(bounded_map.15to21)[3], st_bbox(boundary.sf.4326["CITY"])[3]))
# X_min  <- min(c(st_bbox(bounded_map.15to21)[1], st_bbox(boundary.sf.4326["CITY"])[1]))
# Y_max  <- max(c(st_bbox(bounded_map.15to21)[4], st_bbox(boundary.sf.4326["CITY"])[4]))
# Y_min  <- min(c(st_bbox(bounded_map.15to21)[2], st_bbox(boundary.sf.4326["CITY"])[2]))

X_max  <- max(st_coordinates(bounded_map.15to21)[,1])
X_min  <- min(st_coordinates(bounded_map.15to21)[,1])
Y_max  <- max(st_coordinates(bounded_map.15to21)[,2])
Y_min  <- min(st_coordinates(bounded_map.15to21)[,2])


X_grid  <- seq(X_min, X_max, length.out = 300)
Y_grid  <- seq(Y_min, Y_max, length.out = 300)

spat_grid <- expand.grid(x = X_grid, y = Y_grid)

spat_grid_sf <- st_as_sf(spat_grid, coords = c("x", "y"), crs = 4326)

sf_use_s2(FALSE)

spat_grid_sf <- spat_grid_sf[boundary.sf.4326,]

spat_grid <- as.data.frame(st_coordinates(spat_grid_sf))

names(spat_grid) <- c("x", "y")

sf_use_s2(TRUE)


### PERFORM ACTUAL PREDICTIONS ###

response.preds15   <- as.vector(predict(model15, newdata = spat_grid, type = "response"))
response.preds16   <- as.vector(predict(model16, newdata = spat_grid, type = "response"))
response.preds17   <- as.vector(predict(model17, newdata = spat_grid, type = "response"))
response.preds18   <- as.vector(predict(model18, newdata = spat_grid, type = "response"))
response.preds19   <- as.vector(predict(model19, newdata = spat_grid, type = "response"))
response.preds20   <- as.vector(predict(model20, newdata = spat_grid, type = "response"))
response.preds21   <- as.vector(predict(model21, newdata = spat_grid, type = "response"))
response.preds.all <- as.vector(predict(model.all, newdata = spat_grid, type = "response"))

### BIND PREDICTIONS WITH COORDINATES ###
pred.results15 <- cbind(spat_grid,response.preds15)
pred.results16 <- cbind(spat_grid,response.preds16)
pred.results17 <- cbind(spat_grid,response.preds17)
pred.results18 <- cbind(spat_grid,response.preds18)
pred.results19 <- cbind(spat_grid,response.preds19)
pred.results20 <- cbind(spat_grid,response.preds20)
pred.results21 <- cbind(spat_grid,response.preds21)
pred.results.all <- cbind(spat_grid,response.preds.all)

pred.results15.sf <- st_as_sf(pred.results15, coords = c("x","y"), crs = project_crs)
pred.results16.sf <- st_as_sf(pred.results16, coords = c("x","y"), crs = project_crs)
pred.results17.sf <- st_as_sf(pred.results17, coords = c("x","y"), crs = project_crs)
pred.results18.sf <- st_as_sf(pred.results18, coords = c("x","y"), crs = project_crs)
pred.results19.sf <- st_as_sf(pred.results19, coords = c("x","y"), crs = project_crs)
pred.results20.sf <- st_as_sf(pred.results20, coords = c("x","y"), crs = project_crs)
pred.results21.sf <- st_as_sf(pred.results21, coords = c("x","y"), crs = project_crs)
pred.results.all.sf <- st_as_sf(pred.results.all, coords = c("x","y"), crs = project_crs)


save(model1, model2, model3, 
     model4, model5, model6, 
     model15, model16, model17, 
     model18, model19, model20, 
     model21, model.all,
     file = paste0(fp_data_obj, "boston_regression_objects.RData"))

save(pred.results15, pred.results16, pred.results17, 
     pred.results18, pred.results19, pred.results20, 
     pred.results21, pred.results.all, 
     file = paste0(fp_data_obj, "boston_predictions.RData"))

save(pred.results15.sf, pred.results16.sf, pred.results17.sf, 
     pred.results18.sf, pred.results19.sf, pred.results20.sf, 
     pred.results21.sf, pred.results.all.sf, 
     file = paste0(fp_data_obj, "boston_predictions_sf.RData"))

print("Maximum predicted count:")
max(c(response.preds15, response.preds16, response.preds17, 
      response.preds18, response.preds19, response.preds20, 
      response.preds21))

print("Summary of predicted counts:")
summary(c(response.preds15, response.preds16, response.preds17, 
          response.preds18, response.preds19, response.preds20, 
          response.preds21))



rm(spat_grid, response.preds15, response.preds16, response.preds17, response.preds18, response.preds19, response.preds20, response.preds21, response.preds.all)
```

## Visualize interpolated surfaces (plots)

```{r predicted_plots}
color_scale_max <- max(c(
  max(pred.results15$response.preds15),
  max(pred.results16$response.preds16),
  max(pred.results17$response.preds17),
  max(pred.results18$response.preds18),
  max(pred.results19$response.preds19),
  max(pred.results20$response.preds20),
  max(pred.results21$response.preds21)))

X_max  <- max(st_coordinates(bounded_map.15to21)[,1])
X_min  <- min(st_coordinates(bounded_map.15to21)[,1])
Y_max  <- max(st_coordinates(bounded_map.15to21)[,2])
Y_min  <- min(st_coordinates(bounded_map.15to21)[,2])


basemap.boston <- get_stamenmap(bbox = c(X_min,Y_min,X_max,Y_max),
                    maptype = "terrain-lines", zoom = 13)

cols <- c("#FFFFFF", rev(hcl.colors(16, palette = "YlOrRd")))

### GIF OF OBSERVED
width_in <-  7
height_in <- 5
reso <- 225

png_fp <- file.path(tempdir(), "frame%03d.png")
png(png_fp, width = (width_in * reso), height = (height_in * reso), res = reso)
par(ask = FALSE)


ggmap(basemap.boston) + 
 geom_contour_filled(data = pred.results.all, 
                     aes(x = x, y = y, z = response.preds.all), 
                     breaks = c(0, seq(1, ceiling(max( pred.results.all$response.preds.all)), length.out = 16)), 
                     alpha = .5) + 
 geom_sf(data = boundary.sf.4326, inherit.aes = FALSE, col = "red", alpha = 0.0) +  
 scale_fill_manual(values = cols, drop = FALSE) + 
 labs(title = "Violent Crimes in Boston, 2015-2021", fill = "Total violent \ncrimes expected \nwithin 1/4 mi.") + 
 theme(axis.title.x = element_blank()) + 
 theme(axis.title.y = element_blank()) + 
 coord_sf(crs = st_crs(4326))


ggmap(basemap.boston) + 
 geom_contour_filled(data = pred.results15, 
                     aes(x = x, y = y, z = response.preds15), 
                     breaks = c(0, seq(1, ceiling(color_scale_max), length.out = 16)), 
                     alpha = .5) + 
 geom_sf(data = boundary.sf.4326, inherit.aes = FALSE, col = "red", alpha = 0.0) +  
 scale_fill_manual(values = cols, drop = FALSE) + 
 labs(title = "Violent Crimes in Boston, 2015", fill = "Annual violent \ncrimes expected \nwithin 1/4 mi.") + 
 theme(axis.title.x = element_blank()) + 
 theme(axis.title.y = element_blank()) + 
 coord_sf(crs = st_crs(4326))


ggmap(basemap.boston) + 
 geom_contour_filled(data = pred.results16, 
                     aes(x = x, y = y, z = response.preds16), 
                     breaks = c(0, seq(1, ceiling(color_scale_max), length.out = 16)), 
                     alpha = .5) + 
 geom_sf(data = boundary.sf.4326, inherit.aes = FALSE, col = "red", alpha = 0.0) +  
 scale_fill_manual(values = cols, drop = FALSE) + 
 labs(title = "Violent Crimes in Boston, 2016", fill = "Annual violent \ncrimes expected \nwithin 1/4 mi.") + 
 theme(axis.title.x = element_blank()) + 
 theme(axis.title.y = element_blank()) + 
 coord_sf(crs = st_crs(4326))


ggmap(basemap.boston) + 
 geom_contour_filled(data = pred.results17, 
                     aes(x = x, y = y, z = response.preds17), 
                     breaks = c(0, seq(1, ceiling(color_scale_max), length.out = 16)), 
                     alpha = .5) + 
 geom_sf(data = boundary.sf.4326, inherit.aes = FALSE, col = "red", alpha = 0.0) +  
 scale_fill_manual(values = cols, drop = FALSE) + 
 labs(title = "Violent Crimes in Boston, 2017", fill = "Annual violent \ncrimes expected \nwithin 1/4 mi.") + 
 theme(axis.title.x = element_blank()) + 
 theme(axis.title.y = element_blank()) + 
 coord_sf(crs = st_crs(4326))


ggmap(basemap.boston) + 
 geom_contour_filled(data = pred.results18, 
                     aes(x = x, y = y, z = response.preds18), 
                     breaks = c(0, seq(1, ceiling(color_scale_max), length.out = 16)), 
                     alpha = .5) + 
 geom_sf(data = boundary.sf.4326, inherit.aes = FALSE, col = "red", alpha = 0.0) +  
 scale_fill_manual(values = cols, drop = FALSE) + 
 labs(title = "Violent Crimes in Boston, 2018", fill = "Annual violent \ncrimes expected \nwithin 1/4 mi.") + 
 theme(axis.title.x = element_blank()) + 
 theme(axis.title.y = element_blank()) + 
 coord_sf(crs = st_crs(4326))


ggmap(basemap.boston) + 
 geom_contour_filled(data = pred.results19, 
                     aes(x = x, y = y, z = response.preds19), 
                     breaks = c(0, seq(1, ceiling(color_scale_max), length.out = 16)), 
                     alpha = .5) + 
 geom_sf(data = boundary.sf.4326, inherit.aes = FALSE, col = "red", alpha = 0.0) +  
 scale_fill_manual(values = cols, drop = FALSE) + 
 labs(title = "Violent Crimes in Boston, 2019", fill = "Annual violent \ncrimes expected \nwithin 1/4 mi.") + 
 theme(axis.title.x = element_blank()) + 
 theme(axis.title.y = element_blank()) + 
 coord_sf(crs = st_crs(4326))


ggmap(basemap.boston) + 
 geom_contour_filled(data = pred.results20, 
                     aes(x = x, y = y, z = response.preds20), 
                     breaks = c(0, seq(1, ceiling(color_scale_max), length.out = 16)), 
                     alpha = .5) + 
 geom_sf(data = boundary.sf.4326, inherit.aes = FALSE, col = "red", alpha = 0.0) +  
 scale_fill_manual(values = cols, drop = FALSE) + 
 labs(title = "Violent Crimes in Boston, 2020", fill = "Annual violent \ncrimes expected \nwithin 1/4 mi.") + 
 theme(axis.title.x = element_blank()) + 
 theme(axis.title.y = element_blank()) + 
 coord_sf(crs = st_crs(4326))


ggmap(basemap.boston) + 
 geom_contour_filled(data = pred.results21, 
                     aes(x = x, y = y, z = response.preds21), 
                     breaks = c(0, seq(1, ceiling(color_scale_max), length.out = 16)), 
                     alpha = .5) + 
 geom_sf(data = boundary.sf.4326, inherit.aes = FALSE, col = "red", alpha = 0.0) +  
 scale_fill_manual(values = cols, drop = FALSE) + 
 labs(title = "Violent Crimes in Boston, 2021", fill = "Annual violent \ncrimes expected \nwithin 1/4 mi.") + 
 theme(axis.title.x = element_blank()) + 
 theme(axis.title.y = element_blank()) + 
 coord_sf(crs = st_crs(4326))

ggmap(basemap.boston) + 
 geom_contour_filled(data = pred.results.all, 
                     aes(x = x, y = y, z = response.preds.all), 
                     breaks = c(0, seq(1, ceiling(max( pred.results.all$response.preds.all)), length.out = 16)), 
                     alpha = .5) + 
 geom_sf(data = boundary.sf.4326, inherit.aes = FALSE, col = "red", alpha = 0.0) +  
 scale_fill_manual(values = cols, drop = FALSE) + 
 labs(title = "Violent Crimes in Boston, 2015-2021", fill = "Total violent \ncrimes expected \nwithin 1/4 mi.") + 
 theme(axis.title.x = element_blank()) + 
 theme(axis.title.y = element_blank()) + 
 coord_sf(crs = st_crs(4326))

dev.off()

png_files <- sprintf(png_fp, 1:9)
gif_fp <- paste0(fp_figures, "predicted_violent_crimes.gif")
gifski(png_files, gif_fp, delay = 2)
unlink(png_files)
utils::browseURL(gif_fp)
```

## Spatial satterning in residuals (plots)

```{r resids_spatial_dependence}
# note: A Gaussian GAM with identity link is just an ordinary linear model. In this situation, response residuals, deviance residuals, and pearson residuals are the same.


years <- c("15", "16", "17", "18", "19", "20", "21")


residuals.array <- array(matrix(0, nrow = nrow(modelall.df), ncol = (ncol(modelall.df) + 5)), dim = c(nrow(modelall.df), (ncol(modelall.df) + 5), length(years)))

for (i in years){ 

  resids.response       <- residuals(eval(as.name(paste0("model", i))), type = "response")        # Response residuals: raw residuals (data minus fitted values)



  resids.deviance       <- residuals(eval(as.name(paste0("model", i))), type = "deviance")        # Deviance residuals: residuals defined by model family



  resids.pearson        <- residuals(eval(as.name(paste0("model", i))), type = "pearson")         # Pearson residuals: raw residuals divided by the standard deviation of the data 



  resids.pearson.scaled <- residuals(eval(as.name(paste0("model", i))), type = "scaled.pearson")  # Scaled Pearson residuals: raw residuals divided by the standard deviation of the data 
                                                                                                  # according to the model mean variance relationship and estimated scale parameter.


  resids.working        <- residuals(eval(as.name(paste0("model", i))), type = "working")         # Working residuals: residuals returned from model fitting at convergence


  resids.all            <- cbind(eval(as.name(paste0("model", i, ".df"))), 
                                 resids.response, resids.deviance, resids.pearson, 
                                 resids.pearson.scaled, resids.working)  

  counter               <- as.numeric(which(years == i))
  
  residuals.array[ , , counter] <- as.matrix(resids.all)
  }

resids.15 <- as.data.frame(residuals.array[ , , which(years == "15")])

names(resids.15) <- c("lon", "lat", "crimes", "response.resids", "deviance.resids", 
                      "pearson.resids", "pearson.scaled.resids", "working.resids")

resids.15 <- st_as_sf(resids.15, coords = c("lon", "lat"), crs = st_crs(4326))



resids.16 <- as.data.frame(residuals.array[ , , which(years == "16")])

names(resids.16) <- c("lon", "lat", "crimes", "response.resids", "deviance.resids", 
                      "pearson.resids", "pearson.scaled.resids", "working.resids")

resids.16 <- st_as_sf(resids.16, coords = c("lon", "lat"), crs = st_crs(4326))



resids.17 <- as.data.frame(residuals.array[ , , which(years == "17")])

names(resids.17) <- c("lon", "lat", "crimes", "response.resids", "deviance.resids", 
                      "pearson.resids", "pearson.scaled.resids", "working.resids")

resids.17 <- st_as_sf(resids.17, coords = c("lon", "lat"), crs = st_crs(4326))



resids.18 <- as.data.frame(residuals.array[ , , which(years == "18")])

names(resids.18) <- c("lon", "lat", "crimes", "response.resids", "deviance.resids", 
                      "pearson.resids", "pearson.scaled.resids", "working.resids")

resids.18 <- st_as_sf(resids.18, coords = c("lon", "lat"), crs = st_crs(4326))



resids.19 <- as.data.frame(residuals.array[ , , which(years == "19")])

names(resids.19) <- c("lon", "lat", "crimes", "response.resids", "deviance.resids", 
                      "pearson.resids", "pearson.scaled.resids", "working.resids")

resids.19 <- st_as_sf(resids.19, coords = c("lon", "lat"), crs = st_crs(4326))



resids.20 <- as.data.frame(residuals.array[ , , which(years == "20")])

names(resids.20) <- c("lon", "lat", "crimes", "response.resids", "deviance.resids", 
                      "pearson.resids", "pearson.scaled.resids", "working.resids")

resids.20 <- st_as_sf(resids.20, coords = c("lon", "lat"), crs = st_crs(4326))


resids.21 <- as.data.frame(residuals.array[ , , which(years == "21")])

names(resids.21) <- c("lon", "lat", "crimes", "response.resids", "deviance.resids", 
                      "pearson.resids", "pearson.scaled.resids", "working.resids")

resids.21 <- st_as_sf(resids.21, coords = c("lon", "lat"), crs = st_crs(4326))



width_in <-  7
height_in <- 5
reso <- 225

png_fp <- file.path(tempdir(), "frame%03d.png")
png(png_fp, width = (width_in * reso), height = (height_in * reso), res = reso)
par(ask = FALSE)

ggmap(basemap.boston) + 
        geom_sf(aes(color = pearson.resids), data = resids.15, alpha = 0.75, inherit.aes = F) + 
        scale_color_gradient2("Pearson \nresiduals", low = "blue", mid = "darkgray", high = "red") + 
  ggtitle("Spatial Patterning Within (Pearson) Residuals: 2015") + 
  theme(axis.title = element_blank())   

ggmap(basemap.boston) + 
        geom_sf(aes(color = pearson.resids), data = resids.16, alpha = 0.75, inherit.aes = F) + 
        scale_color_gradient2("Pearson \nresiduals", low = "blue", mid = "darkgray", high = "red") + 
  ggtitle("Spatial Patterning Within (Pearson) Residuals: 2016") + 
  theme(axis.title = element_blank())

ggmap(basemap.boston) + 
        geom_sf(aes(color = pearson.resids), data = resids.17, alpha = 0.75, inherit.aes = F) + 
        scale_color_gradient2("Pearson \nresiduals", low = "blue", mid = "darkgray", high = "red") + 
  ggtitle("Spatial Patterning Within (Pearson) Residuals: 2017") + 
  theme(axis.title = element_blank())

ggmap(basemap.boston) + 
        geom_sf(aes(color = pearson.resids), data = resids.18, alpha = 0.75, inherit.aes = F) + 
        scale_color_gradient2("Pearson \nresiduals", low = "blue", mid = "darkgray", high = "red") + 
  ggtitle("Spatial Patterning Within (Pearson) Residuals: 2018") + 
  theme(axis.title = element_blank())

ggmap(basemap.boston) + 
        geom_sf(aes(color = pearson.resids), data = resids.19, alpha = 0.75, inherit.aes = F) + 
        scale_color_gradient2("Pearson \nresiduals", low = "blue", mid = "darkgray", high = "red") + 
  ggtitle("Spatial Patterning Within (Pearson) Residuals: 2019") + 
  theme(axis.title = element_blank())

ggmap(basemap.boston) + 
        geom_sf(aes(color = pearson.resids), data = resids.20, alpha = 0.75, inherit.aes = F) + 
        scale_color_gradient2("Pearson \nresiduals", low = "blue", mid = "darkgray", high = "red") + 
  ggtitle("Spatial Patterning Within (Pearson) Residuals: 2020") + 
  theme(axis.title = element_blank())

ggmap(basemap.boston) + 
        geom_sf(aes(color = pearson.resids), data = resids.21, alpha = 0.75, inherit.aes = F) + 
        scale_color_gradient2("Pearson \nresiduals", low = "blue", mid = "darkgray", high = "red") + 
  ggtitle("Spatial Patterning Within (Pearson) Residuals: 2021") + 
  theme(axis.title = element_blank())

dev.off()

png_files <- sprintf(png_fp, 1:7)
gif_fp <- paste0(fp_figures, "residual_spatial_patterning.gif")
gifski(png_files, gif_fp, delay = 2)
unlink(png_files)
utils::browseURL(gif_fp)
```

## Construct data objects 

```{r build_data_objs, eval= FALSE}
yr.2015 <- read.csv(paste0(fp_raw_data, "2015_boston.csv"))
yr.2016 <- read.csv(paste0(fp_raw_data, "2016_boston.csv"))
yr.2017 <- read.csv(paste0(fp_raw_data, "2017_boston.csv"))
yr.2018 <- read.csv(paste0(fp_raw_data, "2018_boston.csv"))
yr.2019 <- read.csv(paste0(fp_raw_data, "2019_boston.csv"))
yr.2020 <- read.csv(paste0(fp_raw_data, "2020_boston.csv"))
yr.2021 <- read.csv(paste0(fp_raw_data, "2021_boston.csv"))

#### AGGREGATE DATA ####

yr.15to21 <- rbind(yr.2015, yr.2016, yr.2017, yr.2018, yr.2019, yr.2020, yr.2021)

#### POSIX TIME ####

yr.15to21$POSIX <- as.POSIXlt(strptime(as.character(
                              yr.15to21$OCCURRED_ON_DATE), "%Y-%m-%d %H:%M"))

#### FILTER TO RECORDS WITH SPATIAL COORDS ####
map.15to21 <- yr.15to21[!is.na(yr.15to21$Lat), ]

map.15to21 <- map.15to21[(map.15to21$Lat > -1), ]

#### CREATE SF OBJECT ####
map.15to21 <- st_as_sf(map.15to21, coords = c("Long", "Lat"), crs = 4326)

#### FILTER TO RECORDS WITHIN ZIP CODES ####
zips.sf <- read_sf(dsn = paste0(fp_raw_data, "Zip_Codes"), layer = "ZIP_Codes")

zips.sf <- st_transform(zips.sf, crs = project_crs)

bounded_map.15to21 <- map.15to21[zips.sf, ]

unbounded_map.15to21 <- anti_join(map.15to21, as.data.frame(bounded_map.15to21))

#### ASSIGN ZIP CODES TO RECORDS ####

zip_indices <- sapply(st_intersects(bounded_map.15to21, zips.sf), 
                      function(z) if (length(z) == 0) NA_integer_ else z[1]) # replaces over()

zip_codes  <- zips.sf[zip_indices, "ZIP5"]

st_geometry(zip_codes) <- NULL

names(zip_codes) <- "ZIP5"

zip_codes <- as.data.frame(zip_codes)

bounded_map.15to21$ZIP5 <- NA

bounded_map.15to21$ZIP5 <- zip_codes[, 1]

bounded_map.15to21$ZIP5 <- as.factor(bounded_map.15to21$ZIP5)

#### FILTER TO SERIOUS CRIMES ####
offense_code_groups <- c("Robbery", "Homicide", "HOME INVASION", "Aggravated Assault") 

filter1 <- bounded_map.15to21[bounded_map.15to21$OFFENSE_CODE_GROUP %in% offense_code_groups,]

offense_descriptions_violent <- unique(filter1$OFFENSE_DESCRIPTION)

offense_descriptions_all <- unique(bounded_map.15to21$OFFENSE_DESCRIPTION)

offense_descriptions_all <- offense_descriptions_all[c(21,40,60,87,114,118,126,137,150,189,190,191,192,196,197,202,213,222,230,236,238,250,255,269,275,293)]

offense_descriptions_project <- unique(c(offense_descriptions_violent, offense_descriptions_all))

#### SPLIT BY YEAR ####
boston_crime_spatial_all <- bounded_map.15to21[bounded_map.15to21$OFFENSE_DESCRIPTION %in% offense_descriptions_project,]
boston_crime_spatial_15  <- boston_crime_spatial_all[boston_crime_spatial_all$YEAR == 2015,] 
boston_crime_spatial_16  <- boston_crime_spatial_all[boston_crime_spatial_all$YEAR == 2016,]
boston_crime_spatial_17  <- boston_crime_spatial_all[boston_crime_spatial_all$YEAR == 2017,]
boston_crime_spatial_18  <- boston_crime_spatial_all[boston_crime_spatial_all$YEAR == 2018,]
boston_crime_spatial_19  <- boston_crime_spatial_all[boston_crime_spatial_all$YEAR == 2019,]
boston_crime_spatial_20  <- boston_crime_spatial_all[boston_crime_spatial_all$YEAR == 2020,]
boston_crime_spatial_21  <- boston_crime_spatial_all[boston_crime_spatial_all$YEAR == 2021,]

#### CRIMES WITHOUT SPATIAL INFO ####
boston_crime_missing <- unbounded_map.15to21[unbounded_map.15to21$OFFENSE_DESCRIPTION %in% offense_descriptions_project,]

#### CREATE AGGREGATION GRID AS SF OBJECT ####
cell_size_ft <- 1320

active.grid <- st_make_grid(boundary.sf, cellsize = cell_size_ft, square = F) %>% st_sf() 

active.grid <- st_transform(active.grid, crs = 4326)

active.grid <- dplyr::mutate(active.grid, ID = row_number(),  
               center_x = st_coordinates(st_centroid(active.grid))[, 1],                                      
               center_y = st_coordinates(st_centroid(active.grid))[, 2])

#### *NEW* TRIM HEXAGONS WITH ZERO CRIME ACTIVITY IN 2015-2021 #### 
all_crimes_binned <- st_join(bounded_map.15to21, active.grid)

all_crime_counts <- all_crimes_binned %>% 
  as.data.frame() %>% 
  dplyr::group_by(ID) %>% 
  dplyr::summarize(n = n())

crime_possible_IDs <- unique(all_crime_counts$ID)

crime_possible_IDs <- crime_possible_IDs[1:(length(crime_possible_IDs) - 1)]

active.grid <- active.grid[crime_possible_IDs,]

active.grid <- active.grid[!is.na(active.grid$ID),]

#### BIN OBSERVATIONS #### 

boston_crime_join_all <- st_join(boston_crime_spatial_all, active.grid)
boston_crime_join_15  <- st_join(boston_crime_spatial_15, active.grid)
boston_crime_join_16  <- st_join(boston_crime_spatial_16, active.grid)
boston_crime_join_17  <- st_join(boston_crime_spatial_17, active.grid)
boston_crime_join_18  <- st_join(boston_crime_spatial_18, active.grid)
boston_crime_join_19  <- st_join(boston_crime_spatial_19, active.grid)
boston_crime_join_20  <- st_join(boston_crime_spatial_20, active.grid)
boston_crime_join_21  <- st_join(boston_crime_spatial_21, active.grid)

crime_counts_15 <- boston_crime_join_15 %>% 
  as.data.frame() %>% 
  dplyr::group_by(ID) %>% 
  dplyr::summarize(n = n())

crime_counts_16 <- boston_crime_join_16 %>% 
  as.data.frame() %>% 
  dplyr::group_by(ID) %>% 
  dplyr::summarize(n = n())

crime_counts_17 <- boston_crime_join_17 %>% 
  as.data.frame() %>% 
  dplyr::group_by(ID) %>% 
  dplyr::summarize(n = n())

crime_counts_18 <- boston_crime_join_18 %>% 
  as.data.frame() %>% 
  dplyr::group_by(ID) %>% 
  dplyr::summarize(n = n())

crime_counts_19 <- boston_crime_join_19 %>% 
  as.data.frame() %>% 
  dplyr::group_by(ID) %>% 
  dplyr::summarize(n = n())

crime_counts_20 <- boston_crime_join_20 %>% 
  as.data.frame() %>% 
  dplyr::group_by(ID) %>% 
  dplyr::summarize(n = n())

crime_counts_21 <- boston_crime_join_21 %>% 
  as.data.frame() %>% 
  dplyr::group_by(ID) %>% 
  dplyr::summarize(n = n())

crime_counts_all <- boston_crime_join_all %>% 
  as.data.frame() %>% 
  dplyr::group_by(ID) %>% 
  dplyr::summarize(n = n())

project.grid.15  <- left_join(active.grid, crime_counts_15, by = "ID")
project.grid.16  <- left_join(active.grid, crime_counts_16, by = "ID")
project.grid.17  <- left_join(active.grid, crime_counts_17, by = "ID")
project.grid.18  <- left_join(active.grid, crime_counts_18, by = "ID")
project.grid.19  <- left_join(active.grid, crime_counts_19, by = "ID")
project.grid.20  <- left_join(active.grid, crime_counts_20, by = "ID")
project.grid.21  <- left_join(active.grid, crime_counts_21, by = "ID")
project.grid.all <- left_join(active.grid, crime_counts_all, by = "ID")

for (i in 1:nrow(project.grid.15)){
  if (is.na(st_drop_geometry(project.grid.15[i, "n"]))) {project.grid.15[i, "n"] <- 0}
  
  else {project.grid.15[i, "n"] <- st_drop_geometry(project.grid.15[i, "n"])}
  }

for (i in 1:nrow(project.grid.16)){
  if (is.na(st_drop_geometry(project.grid.16[i, "n"]))) {project.grid.16[i, "n"] <- 0}
  
  else {project.grid.16[i, "n"] <- st_drop_geometry(project.grid.16[i, "n"])}
  }

for (i in 1:nrow(project.grid.17)){
  if (is.na(st_drop_geometry(project.grid.17[i, "n"]))) {project.grid.17[i, "n"] <- 0}
  
  else {project.grid.17[i, "n"] <- st_drop_geometry(project.grid.17[i, "n"])}
  }

for (i in 1:nrow(project.grid.18)){
  if (is.na(st_drop_geometry(project.grid.18[i, "n"]))) {project.grid.18[i, "n"] <- 0}
  
  else {project.grid.18[i, "n"] <- st_drop_geometry(project.grid.18[i, "n"])}
  }

for (i in 1:nrow(project.grid.19)){
  if (is.na(st_drop_geometry(project.grid.19[i, "n"]))) {project.grid.19[i, "n"] <- 0}
  
  else {project.grid.19[i, "n"] <- st_drop_geometry(project.grid.19[i, "n"])}
  }

for (i in 1:nrow(project.grid.20)){
  if (is.na(st_drop_geometry(project.grid.20[i, "n"]))) {project.grid.20[i, "n"] <- 0}
  
  else {project.grid.20[i, "n"] <- st_drop_geometry(project.grid.20[i, "n"])}
  }

for (i in 1:nrow(project.grid.21)){
  if (is.na(st_drop_geometry(project.grid.21[i, "n"]))) {project.grid.21[i, "n"] <- 0}
  
  else {project.grid.21[i, "n"] <- st_drop_geometry(project.grid.21[i, "n"])}
  }

for (i in 1:nrow(project.grid.all)){
  if (is.na(st_drop_geometry(project.grid.all[i, "n"]))) {project.grid.all[i, "n"] <- 0}
  
  else {project.grid.all[i, "n"] <- st_drop_geometry(project.grid.all[i, "n"])}
  }


save(project.grid.all, project.grid.15, project.grid.16, 
     project.grid.17, project.grid.18, project.grid.19, 
     project.grid.20, project.grid.21, boston_crime_missing, 
     boston_crime_spatial_all, project_crs, yr.15to21, bounded_map.15to21, 
     boundary.sf, boundary.sf.4326, 
     file = paste0(fp_data_obj, "boston_spatial_project_r1.RData"))


rm(yr.2015, yr.2016, yr.2017, yr.2018, yr.2019, yr.2020, yr.2021, filter1, zip_codes, offense_descriptions_violent, offense_descriptions_all, crime_counts_15, crime_counts_16, crime_counts_17, crime_counts_18, crime_counts_19, crime_counts_20, crime_counts_21, crime_counts_all, map.15to21, unbounded_map.15to21, crime_possible_IDs, offense_code_groups, zip_indices, all_crimes_binned, all_crime_counts, zips.sf, boston_crime_join_all, boston_crime_join_15, boston_crime_join_16, boston_crime_join_17, boston_crime_join_18, boston_crime_join_19, boston_crime_join_20, boston_crime_join_21, boston_crime_spatial_all, boston_crime_spatial_15, boston_crime_spatial_16, boston_crime_spatial_17, boston_crime_spatial_18, boston_crime_spatial_19, boston_crime_spatial_20, boston_crime_spatial_21, active.grid)
```
